
# Import Packages
```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

# Installs pacman package if needed
if (!require("pacman")) install.packages("pacman")

# load packages
pacman::p_load(
    pacman,     # package installer
    installr,   # software installer
    rio,        # data importing
    dplyr,      # data manipulation
    tidyr,      # data manipulation (gather etc.)
    ggplot2,    # ploting
    stringr,    # string formatting
    scales,     # plot scales and color etc. 
    viridis,    # color palette
    hrbrthemes, # plot themes
    plotly,     # interactive plot
    treemap,    # tree plot
    htmlwidgets # save interactive plot into html
    )

# This is needed if packages are giving errors for incompatible binary distribution
# install.Rtools() 

devtools::install_github("timelyportfolio/d3treeR")
library(d3treeR) # interactive tree plot

# load custom functions
source("./functions/general_functions.R")

```

# Specify Paths
```{r}
# inputs
dataset_path <- "./Data/"
function_path <- "./Functions/"

# outputs
itrt_plot_path <- "./Output/InteractivePlot/"
sttc_plot_path <- "./Output/StaticPlot/"
out_data_path <- "./Output/Data/"

```

# Import Dataset
```{r}

daily_covid <- import(
    paste0(dataset_path, "worldometer_coronavirus_daily_data.csv")
    )

summary_covid <- import(
    paste0(dataset_path, "worldometer_coronavirus_summary_data.csv")
    )

head(daily_covid)
head(summary_covid)

```
# Data Cleaning
```{r}

# daily_covid[is.na(d)] <- 0

```

# Data Manipulation and visualization
## Generic Questions and Answers
### Question 1 
What is the overview of covid cases during that period of time?
```{r}

## global percentage of death, active case and recovered ##

# sum vertically
categories <- c("total_deaths", "total_recovered", "active_cases")
category <- str_replace_all(categories, pattern =  "_", replacement = " ")
category <- str_to_title(category)

data <- 
    summary_covid[, categories] %>%
    colSums(na.rm = T)
data <- data.frame(
  category=category,
  count=data
)

data$prettyCount <- prettyNum(data$count, big.mark = ",", scientific = F)

# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# Compute display percentages
data$prettyFraction <- percent(data$fraction)

# Make the plot
q1 <- 
    ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
    geom_rect() +
    geom_text( 
        x=4.3, 
        aes(y=labelPosition, label=prettyCount, color=category, fontface="bold"), 
        size=3.5
    ) + # x here controls label position (inner / outer)
    geom_text(
        x=3.5, 
        aes(y=labelPosition, label=prettyFraction, fontface="bold"), 
        color="white",
        size=4
    ) +
    scale_fill_brewer(palette="Set2") +
    scale_color_brewer(palette="Set2") +
    coord_polar(theta="y") +
    xlim(c(2, 4)) +
    theme_void() +
    annotate(
        geom = "text", 
        x = 2, 
        y = 0, 
        colour = "#eba834",
        label = paste0(
            "Total Cases\n", 
            prettyNum(sum(data$count), big.mark = ",", scientific = F)
        )
    )

ggsave(paste0(sttc_plot_path, "q1_pie.png"))
q1

```

### Question 2
What are the general scale of cases within different continent or country?
```{r}
## comparison of cases between different continent ##

q2 <- 
    summary_covid %>% # data
    select(country:active_cases) %>%
    group_by(continent) %>% # group_by 
    filter(total_confirmed > quantile(total_confirmed, 0.7)) %>% # removing small cases
    ungroup() %>% 
    group_by(continent, country) %>% 
    # turning 3 columes into sub sub group (wide to long conversion)
    gather(category, count, total_recovered, active_cases, total_deaths, factor_key=T) %>% 
    ungroup() %>% 
    mutate(category = factor(category, labels = c("Recovered", "Active Cases", "Deaths"))) %>% 
    treemap( index=c("continent","country","category"),
             vSize="count",
             type="index",
             palette = "Set2",
             title = "Group by continent top 70 percentile confirmed cases",
             align.labels=list(
                 c("center", "center"),
                 c("left", "top"),
                 c("left", "bottom")
             )
        )

itrt_q2 <- d3tree2( q2 ,  rootname = "Group by continent top 70 percentile confirmed cases" )

saveWidget(itrt_q2, file = paste0(itrt_plot_path, "q2"))
itrt_q2

```

### Question 3
What is the spreading trend in that period of time?

```{r}

# overview of accumulated cases vs date for all the country
# 

```

### Question 4
Which country has the most cases?

```{r}

# ranking of cases for the top n countries
# hist

```

## Question 5
How active cases have change during that period of time for each country?
```{r}
# video hist
```


## More in depth Questions
### Question 1
We can easily tell from previous plots that most of the cases are from big countries. 
What is the relation between population and Covid cases?
```{r fig.align="center", fig.width = 12, fig.height = 6}

# we can easily tell from previous plots that most of the cases are from big
# countries. Now, I'm curious about the relation between population & Covid cases

q11 <- summary_covid %>%
    
    # Reorder countries to having big bubbles at the back
    arrange(desc(population)) %>%
    
    # prepare text for tooltip
    mutate(text = 
               paste0(
                   "Country: ", country, 
                   "\nPopulation: ", commaNum(population), 
                   "\nTotal Cases:\t", commaNum(total_confirmed), 
                   "\nTotal Tests\t", commaNum(total_tests)
               )
           ) %>%
    
    ggplot( 
        aes(
            x = population, 
            y = total_confirmed, 
            fill = continent, 
            size = total_tests, 
            text = text
        )
    ) +
    geom_point(alpha=0.5, color = "black", shape = 21, na.rm = T) +
    scale_x_log10(
        labels = unit_format(unit = "M", scale = 1e-6),
        breaks = 1e+3 * 10^(seq(0,20,2)),
    ) +
    scale_y_log10(
        labels = unit_format(unit = "M", scale = 1e-6),
        breaks = 10^(seq(1,21,2)),
    ) +
    scale_size(range = c(2, 25), name="Total Tests (M), Size") +
    labs(fill = 'Continent, Color') +
    scale_fill_viridis(discrete=T, option = "D") +
    coord_cartesian(clip = "off") +
    ylab("Covid Cases (M), log10(n)") +
    xlab("Population (M), log10(n)") +
    theme_bw()

ggsave(paste0(sttc_plot_path, "q11_bubble.png"))

# turn interactive ggplot with plotly and save it
itrt_q11 <- ggplotly(q11, tooltip="text")
saveWidget(itrt_q11, file = paste0(itrt_plot_path, "q11"))

q11
itrt_q11

```
### Question 2
Which country did well in this pandemic war?
```{r}

```

### Question 3
can weather, geometric position affect the spread of covid?

### Question 4
an even more in depth question, we've seen that population definitely affect how
covid spread, let's check and see why social distancing is needed.

# clean up
```{r}

pacman::p_unload(all)

```

